---
title: "Science Curiosity Study 1 Data"
output: pdf_document
always_allow_html: true
---

```{r mess-warn, include=FALSE}
knitr::opts_chunk$set(message=F, warning=F)
```

```{r lib, echo=F}
library(rio)
library(here)
library(janitor)
library(reshape2)

library(tidyverse)
library(ggplot2)
library(lme4)

library(knitr)
library(Hmisc)
library(viridis)
library(kableExtra)
```

```{r data, echo=F}
dfd1 <- import(here("analysis-data", "cleaned-data0815.csv"))
dfd1_dt <- import(here("analysis-data", "cleaned-data0815-exdt.csv"))
dfd1_flr <- import(here("analysis-data", "cleaned-data0815-exflr.csv"))

dfd1_lc <- dfd1 %>% 
  filter(conservatism_f != 4 | con != 0)
dfd1_lcdt <- dfd1_dt %>% 
  filter(conservatism_f != 4 | con != 0)
dfd1_lcflr <- dfd1_flr %>% 
  filter(conservatism_f != 4 | con != 0)
```

```{r func-se, echo=F}
se_f <- function(x) {
  
  if(length(x) > 0) {
    sd(x, na.rm = T) / sqrt(length(x))
  }
  else {
    999
  }
  
}
```

```{r se-test, echo=F}
# testing function:

# se_func(cylll)

# dfd2_lc %>% 
#   select(sc_scoredz, rav_scoredz) %>% 
#   psych::describe()
# 
# dfd2_lc$rav_scoredz %>% 
#   unique()
# 
# dfd2_lc %>% 
#   group_by(rav_scoredz) %>% 
#   summarise(n = n())
# 
# dfd2_lc %>%
#   filter(is.na(rav_scoredz) == T) %>% 
#   nrow()
# 
# dfd2_lc %>%
#   filter(is.na(rpm_3cat) == T) %>% 
#   nrow()
# 
# dfd2_lc %>% 
#   filter(rav_scoredz >= 2) %>% 
#   nrow()
```

```{r upper-tri, echo=F}
get_upper_tri_list <- function(cormat){
    
  r <- cormat[1]
  n <- cormat[2]
  p <- cormat[3]
  
  r[lower.tri(r)] <- NA
  p[lower.tri(p)] <- NA
  
  cormat_u <- list(r, n, p)
    
    return(cormat_u)
}

get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
```

# Demographics
```{r Ss, echo=F}
n_total <- dfd1 %>% 
  nrow()

n_total_lc <- dfd1_lc %>% 
  nrow()

n_by_con <- dfd1 %>% 
  group_by(con) %>% 
  summarise(n = n(),
            percent = round(100 * (n()/nrow(dfd1)), 2))

n_by_con_lc <- dfd1_lc %>% 
  group_by(con) %>% 
  summarise(n = n(),
            percent = round(100 * (n()/nrow(dfd1_lc)), 2))

df_age <- dfd1_lc %>% 
  select(dem_age_b) %>% 
  psych::describe()

m_age <- df_age$mean %>% 
  round(2)
min_age <- df_age$min %>% 
  round(2)
max_age <- df_age$max %>% 
  round(2)

df_gender <- dfd1_lc %>% 
  group_by(gender_b) %>% 
  summarise(n = n(),
            percent = round(100 *(n()/nrow(dfd1)),2))

df_race <- dfd1_lc %>% 
  group_by(race_b) %>% 
  summarise(n = n(),
            percent = round(100 *(n()/nrow(dfd1)),2))
```

N$_{total}$ = `r n_total`: N$_{con}$ = `r n_by_con[3,2]`, N$_{lib}$ = `r n_by_con[1,2]`, N$_{mod}$ = `r n_by_con[2,2]`

**N$_{final}$ = `r n_total_lc`**

- **Political ideology:** `r n_by_con_lc[2,3]`% conservative, `r n_by_con_lc[1,3]`% liberal
- **Age:** Mean$_{age}$ = `r m_age` (range: `r min_age`-`r max_age`)
- **Gender:** `r df_gender[3,3]`% women, `r df_gender[1,3]`% men, `r df_gender[2,3]`% non-binary folks, `r df_gender[4,3]`% preferred not to answer
- **Race:** `r df_race[1,3]`% White, `r df_race[2,3]`% Black, `r df_race[3,3]`% Hispanic, `r df_race[4,3]`% Asian

# Descriptives
## Summary
```{r basic-desc-tab, echo=F}
dfd1_lc %>% 
  mutate(RPM = rav_scored,
         SciCuriosity = sc_scored,
         ConArctic = arctic20,
         ConTemp = temp20,
         LibOzone = ozone20,
         LibAir = airqual20,
         ConFillCO2 = co220,
         ConFillIce = icesheets20,
         FillBacteria = diatho20,
         FillQuake = quake20) %>% 
  select(SciCuriosity, RPM, ConArctic, ConTemp, LibOzone, LibAir, ConFillIce, FillBacteria, FillQuake) %>% 
  psych::describe() %>% 
  mutate(Min = min,
         Max = max,
         Mean = mean,
         SE = se,
         Median = median) %>% 
  as.data.frame() %>% 
  select(Mean, SE, Min, Max, Median) %>% 
  kable(digits = 2, align = "c") %>% 
  kable_styling(full_width = F, position = "left") %>% 
  column_spec(column = 1, bold = T) %>% 
  column_spec(column = c(2,4,6),background = "lightgray") %>% 
  column_spec(column = 1, border_right = T)
```

```{r add-desc-names, echo=F}
dfd1_lc <- dfd1_lc %>% 
  mutate(RPM = rav_scored,
         SciCuriosity = sc_scored,
         ConArctic = arctic20,
         ConTemp = temp20,
         LibOzone = ozone20,
         LibAir = airqual20,
         ConFillCO2 = co220,
         ConFillIce = icesheets20,
         FillBacteria = diatho20,
         FillQuake = quake20,
         Con = con)
```

## Correlations
```{r cor-map-prep, echo=F}
dfp_corr <- dfd1_lc %>% 
  select(conservatism_f, sc_scored, rav_scored, arctic20, temp20, ozone20, airqual20, co220, icesheets20, diatho20, quake20) %>% 
  mutate(Con = conservatism_f,
         SciCur = sc_scored,
         RPM = rav_scored,
         ConArctic = arctic20,
         ConTemp = temp20,
         LibOzone = ozone20,
         LibAirQual = airqual20,
         ConFillCO2 = co220,
         ConFillIce = icesheets20) %>% 
  select(-contains("20"), -contains("scored"), -contains("_f"))
```

```{r cor-mat-prep2, echo=F}
df_cor <- rcorr(as.matrix(dfp_corr))[[1]] 

# %>% 
#   as.data.frame()

df_pcor <- rcorr(as.matrix(dfp_corr))[[3]] %>% 
  get_upper_tri() %>% 
  melt() %>% 
  mutate(p05 = case_when(
    value < 0.1 & value >= 0.05 ~ 1,
    value < 0.05 & value >= 0.01 ~ 2,
    value < 0.01 & value >= 0.001 ~ 3,
    value < 0.001 ~ 4,
    TRUE ~ 999
  ))

ps <- df_pcor$p05

```

```{r cor-mat1, echo=F}
df_cor %>% 
  get_upper_tri() %>% 
  melt() %>% 
  ggplot(aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = factor(ps)), color = "black") +
  scale_fill_manual(values = c("#66CC00", "#FFFF66", "#FF9933", "#FF3300", "white"), name = "p values",
                    labels = c(
                      "0.1 > p > 0.05",
                      "0.05 > p > 0.01",
                      "0.01 > p > 0.001",
                      "p > 0.001",
                      "ignore :~)"
                    )) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25, vjust = 1, 
    size = 13, hjust = 1)) +
  coord_fixed() +
  geom_text(aes(Var1, Var2, label = round(value, 2)), size = 3) +
  theme(
    axis.text.y = element_text(size = 13),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
  panel.background = element_blank(),
  panel.grid.major = element_blank()) +
  labs(caption = "Conservatism (Con): Extremely liberal (1) to Extremely conservative (7).")
```

```{r heat-map, echo=F, eval=F}
cor(dfp_corr) %>% 
  get_upper_tri() %>% 
  melt() %>% 
  ggplot(aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_viridis(limits = c(-1,1), discrete=FALSE, option = "A",
                     name="Pearson\nCorrelation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25, vjust = 1, 
    size = 13, hjust = 1)) +
  coord_fixed() +
  geom_text(aes(Var1, Var2, label = round(value, 2)), color = "white", size = 3) +
  theme(
    axis.text.y = element_text(size = 13),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
  panel.background = element_blank(),
  panel.grid.major = element_blank()) +
  labs(caption = "Conservatism (Con): Extremely liberal (1) to Extremely conservative (7).")
```

## Science curiosity and RPM

**Grouped by ideology**

```{r desc-table-ideo2, echo=F}
desc_scrpm_con <- dfd1_lc %>% 
  filter(Con == 1) %>% 
  select(SciCuriosity, RPM) %>%
  psych::describe()%>% 
  mutate(Min_c = min,
         Max_c = max,
         Mean_c = mean,
         SE_c = se) %>% 
  as.data.frame() %>% 
  select(Mean_c, SE_c, Min_c, Max_c)

desc_scrpm_lib <- dfd1_lc %>% 
  filter(Con == -1) %>% 
  select(SciCuriosity, RPM) %>%
  psych::describe()%>% 
  mutate(Min_l = min,
         Max_l = max,
         Mean_l = mean,
         SE_l = se) %>% 
  as.data.frame() %>% 
  select(Mean_l, SE_l, Min_l, Max_l)

cbind(desc_scrpm_con, desc_scrpm_lib) %>% 
  select(contains("_c"), contains("_l")) %>% 
  kable(digits = 2, align = "c",
        col.names = c(
          rep(c("Mean", "SE", "Min", "Max"), 2)
        )) %>% 
  add_header_above(c(" " = 1, "Conservative" = 4, "Liberal" = 4), line_sep = 25) %>% 
  kable_styling(full_width = F, position = "left")%>% 
  column_spec(column = 1, bold = T) %>% 
  column_spec(column = c(2,4,6,8), background = "lightgray") %>% 
  column_spec(column = c(1,5), border_right = T)
```

```{r prep-for-plot, echo=F}
desc_long <- dfd1_lc %>% 
  pivot_longer(cols = c(SciCuriosity, RPM, ConArctic, ConTemp, LibOzone, LibAir, ConFillIce, FillBacteria, FillQuake), names_to = "Variable", values_to = "Score")

desc_long_mr <- desc_long %>% 
  filter(Variable != "SciCuriosity" & Variable != "RPM") 
```

```{r cont-plot, echo=F}
desc_long_cont <- desc_long %>% 
  filter(Variable == "SciCuriosity" | Variable == "RPM")

# desc_long_cont %>% 
#   select(Score) %>% 
#   unique()

desc_long_cont %>% 
  ggplot(., aes(x = as.factor(Con), y = Score, fill = as.factor(Con))) + 
  stat_summary(geom = "bar", fun = "mean", position = "dodge") +
  facet_wrap(~Variable, scales = "free_y") +
  scale_x_discrete(labels = c("Lib", "Con")) +
  scale_fill_manual(values = c("deepskyblue3", "brown3")) +
  theme_bw() +
  labs(title = "Average Cognitive Ability and Science Curiosity",
       y = "Average Score",
       x = "Ideology") +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.position = "none",
        plot.title = element_text(size = 16, face = "bold", hjust = .5))
```

## Motivated reasoning items

**Grouped by ideology**

```{r desc-table-ideo, echo=F}
desc_mr_con <- dfd1_lc %>% 
  filter(Con == 1) %>% 
  select(ConArctic, ConTemp, LibOzone, LibAir, ConFillIce, FillBacteria, FillQuake) %>% 
  psych::describe()%>% 
  mutate(Min_c = min,
         Max_c = max,
         Mean_c = mean,
         SE_c = se) %>% 
  as.data.frame() %>% 
  select(Mean_c, SE_c, Min_c, Max_c)

desc_mr_lib <- dfd1_lc %>% 
  filter(Con == -1) %>% 
  select(ConArctic, ConTemp, LibOzone, LibAir, ConFillIce, FillBacteria, FillQuake) %>%
  psych::describe() %>% 
  mutate(Min_l = min,
         Max_l = max,
         Mean_l = mean,
         SE_l = se) %>% 
  as.data.frame() %>% 
  select(Mean_l, SE_l, Min_l, Max_l)

cbind(desc_mr_con, desc_mr_lib) %>% 
  select(contains("_c"), contains("_l")) %>% 
  kable(digits = 2, align = "c",
        col.names = c(
          rep(c("Mean", "SE", "Min", "Max"), 2)
        )) %>% 
  add_header_above(c(" " = 1, "Conservative" = 4, "Liberal" = 4), line_sep = 25) %>% 
  kable_styling(full_width = F, position = "left")%>% 
  column_spec(column = 1, bold = T) %>% 
  column_spec(column = c(2,4,6,8), background = "lightgray") %>% 
  column_spec(column = c(1,5), border_right = T)
```

```{r eb-ideo-plot, echo=F}
desc_long_mr %>% 
  ggplot(., aes(x = Variable, y = Score, fill = as.factor(Con))) + 
  stat_summary(geom = "bar", fun = "mean", position = "dodge") +
  scale_fill_manual(values = c("deepskyblue3", "brown3"), name = "Ideology", labels = c("Lib", "Con")) +
  theme_bw() +
  scale_y_continuous(limits = c(-3, 1), breaks = seq(from  = -3, to = 1, by = .5)) +  
  labs(title = "Average Scores on Motivated Reasoning Items",
       subtitle = "Higher scores indicate more endpoint bias.",
       y = "Average Score",
       x = "Item") +
  theme(axis.text.x = element_text(angle = 45, size = 12, vjust = .6),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        plot.title = element_text(size = 16, face = "bold", hjust = .5),
        plot.subtitle = element_text(size = 12, hjust = .13))
```

# "3-way int." means
```{r add-cong-var, echo=F}
dfd2_lc <- dfd1_lc %>% 
  mutate(ArcticCong = case_when(
    Con == 1 ~ 1,
    Con == -1 ~ 0,
    TRUE ~ 999),
    TempCong = case_when(
      Con == 1 ~ 1,
      Con == -1 ~ 0,
      TRUE ~ 999),
    OzoneCong = case_when(
      Con == 1 ~ 0,
      Con == -1 ~ 1,
      TRUE ~ 999),
    AirCong = case_when(
      Con == 1 ~ 0,
      Con == -1 ~ 1,
      TRUE ~ 999)
         )
```

```{r scrpm-hilow, echo=F}
# dfd2_lc <- dfd2_lc %>% 
#   mutate(scicur_3cat = case_when(
#     sc_scoredz <= -2 ~ -2,
#     sc_scoredz > -2 & sc_scoredz < 0 ~ -1,
#     sc_scoredz == 0 ~ 0,
#     sc_scoredz > 0 & sc_scoredz < 2 ~ 1,
#     sc_scoredz >= 2 ~ 2
#   ),
#   rpm_3cat = case_when(
#     rav_scoredz <= -2 ~ -2,
#     rav_scoredz > -2 & sc_scoredz < 0 ~ -1,
#     rav_scoredz == 0 ~ 0,
#     rav_scoredz > 0 & sc_scoredz < 2 ~ 1,
#     rav_scoredz >= 2 ~ 2
#   )
#   )

dfd2_lc <- dfd2_lc %>% 
  mutate(sc3 = case_when(
    sc_scoredz < 0 ~ "Low",
    sc_scoredz == 0 ~ "Avg",
    sc_scoredz > 0 ~ "High"
  ),
  rpm3 = case_when(
    rav_scoredz < 0 ~ "Low",
    rav_scoredz == 0 ~ "Avg",
    rav_scoredz > 0 ~ "High"
  ),
  ideo2 = case_when(
    con == 1 ~ "Con",
    con == -1 ~ "Lib"
  )
  )
```

```{r 3way-means, echo = F}
int_means <- dfd2_lc %>% 
  group_by(sc3, rpm3, ideo2) %>% 
  summarise_at(vars("ConArctic", "ConTemp", "LibOzone", "LibAir"), funs(mean, se_f, length))

int_means_tab <- int_means %>% 
  select(-contains("length")) %>% 
  select(sc3, rpm3, ideo2, contains("Arctic"), contains("Temp"), contains("Ozone"), contains("Air"))

int_means_tab %>% 
  select(ideo2, sc3, rpm3, everything()) %>% 
  arrange(ideo2, sc3) %>% 
  kable(digits = 2,
        align = "c",
        col.names = c("", "SciCur", "RPM", "Con1_M", "Con1_SE", "Con2_M", "Con2_SE", "Lib1_M", "Lib1_SE", "Lib2_M", "Lib2_SE")) %>% 
  add_header_above(c(" " = 3, "Conservative" = 4, "Liberal" = 4), line_sep = 25) %>%
  kable_styling(position = "left") %>% 
  row_spec(row = c(1, 2, 5, 6), background = "lightgray") %>% 
  column_spec(column = 1, bold = T, color = "white", background = c("maroon", "maroon", "maroon", "maroon", "royalblue", "royalblue", "royalblue", "royalblue")) %>% 
  column_spec(colum = c(3, 5, 7, 9), border_right = T) %>% 
  footnote(general = "* M = mean, SE = standard error \n * Item names: Con1 = arctic sea ice, Con2 = global temperature index, Lib1 = ozone layer hole, Lib2 = US air quality")
```

# Conservative #1: Arctic
```{r echo=F}
int_means_tab %>% 
  ggplot(aes(x = sc3, y = ConArctic_mean, fill = ideo2)) +
  geom_bar(stat = "identity", position = "dodge")
```

```{r}
arcmod1 <- lm(ConArctic ~ 1, data = dfd2_lc)
summary(arcmod1)

arcmod2 <- lm(ConArctic ~ rav_scoredz, data = dfd2_lc)
summary(arcmod2)

anova(arcmod1, arcmod2)

arcmod3 <- lm(ConArctic ~ sc_scoredz + rav_scoredz, data = dfd2_lc)
summary(arcmod3)

anova(arcmod1, arcmod3)

arcmod4 <- lm(ConArctic ~ ArcticCong + sc_scoredz + rav_scoredz, data = dfd2_lc)
summary(arcmod4)

anova(arcmod1, arcmod4)
```

# Conservative #2: Temp
```{r echo=F}
int_means_tab %>% 
  ggplot(aes(x = sc3, y = ConTemp_mean, fill = ideo2)) +
  geom_bar(stat = "identity", position = "dodge")
```

```{r}
tempmod1 <- lm(ConTemp ~ 1, data = dfd2_lc)
summary(tempmod1)

tempmod2 <- lm(ConTemp ~ rav_scoredz, data = dfd2_lc)
summary(tempmod2)

anova(tempmod1, tempmod2)

tempmod3 <- lm(ConTemp ~ sc_scoredz + rav_scoredz, data = dfd2_lc)
summary(tempmod3)

anova(tempmod1, tempmod3)

tempmod4 <- lm(ConTemp ~ TempCong + sc_scoredz + rav_scoredz, data = dfd2_lc)
summary(tempmod4)

anova(tempmod3, tempmod4)
```

# Liberal #1: Ozone
```{r echo=F}
int_means_tab %>% 
  ggplot(aes(x = sc3, y = LibOzone_mean, fill = ideo2)) +
  geom_bar(stat = "identity", position = "dodge")
```

```{r}
ozmod1 <- lm(LibOzone ~ 1, data = dfd2_lc)
summary(ozmod1)

ozmod2 <- lm(LibOzone ~ rav_scoredz, data = dfd2_lc)
summary(ozmod2)

anova(ozmod1, ozmod2)

ozmod3 <- lm(LibOzone ~ sc_scoredz + rav_scoredz, data = dfd2_lc)
summary(ozmod3)

anova(ozmod1, ozmod3)

ozmod4 <- lm(LibOzone ~ OzoneCong + sc_scoredz + rav_scoredz, data = dfd2_lc)
summary(ozmod4)

anova(ozmod1, ozmod4)
```

```{r 3way-plot, echo=F}
# +
#   geom_errorbar(aes(ymin = LibAir_rmean-LibAir_rse, ymax = LibAir_rmean+LibAir_rse), width = .2, position = position_dodge(.9))
```

# Liberal #2: Air Quality
```{r echo=F}
int_means_tab %>% 
  ggplot(aes(x = sc3, y = LibAir_mean, fill = ideo2)) +
  geom_bar(stat = "identity", position = "dodge") 
```

```{r}
airmod1 <- lm(LibAir ~ 1, data = dfd2_lc)
summary(airmod1)

airmod2 <- lm(LibAir ~ rav_scoredz, data = dfd2_lc)
summary(airmod2)

anova(airmod1, airmod2)

airmod3 <- lm(LibAir ~ sc_scoredz + rav_scoredz, data = dfd2_lc)
summary(airmod3)

anova(airmod1, airmod3)

airmod4 <- lm(LibAir ~ AirCong + sc_scoredz + rav_scoredz, data = dfd2_lc)
summary(airmod4)

anova(airmod1, airmod4)
```

```{r mlm-attempt, echo=F, eval=F}
# rcorr(dfd1_lc$sc_scored, dfd1_lc$rav_scored)

sc_1 <- lm(sc_scored ~ 1, dfd1_lc)
summary(sc_1)
scrpm <- lm(sc_scored ~ 1 + rav_scored, dfd1_lc)
summary(scrpm)
anova(sc_1, scrpm)
```