---
title: "prep-raw-data"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
---

# Libraries

Load necessary packages.

```{r load-libs, message=F, warning=F}
library(rio)
library(here)
library(tidyverse)
library(janitor)
```

# Baseline Data

Import baseline data.

```{r import-b-data}
# mturk
dfm_bp <- import(here("data", "baselinep-data.csv")) %>% 
  clean_names() %>% 
  rename(., "id" = "amazon_identifier") %>% 
  mutate(batch = "basep") # import HIT 1

dfm_b1 <- import(here("data", "baseline1-data.csv")) %>% 
  clean_names() %>% 
  rename(., "id" = "amazon_identifier") %>% 
  mutate(batch = "base1") # import HIT 2

# qualtrics
dfq_b <- import(here("data", "qual-baseline-data.csv")) %>% 
  clean_names() # import
```

## MTurk

Join separate HITs into one df.

```{r bind-b}
dfm_b <- rbind(dfm_bp, dfm_b1) # bind HIT dfs
```

Identify duplicates and approved responses.

```{r id-dup/app-b}
dfm_b_dup <- dfm_b %>% 
  group_by(id) %>% 
  filter(n() > 1) # identify duplicates

dfm_b_app <- dfm_b %>% 
  filter(approval_status == "Approved") # identify approved HITs
```

### Save dfs

Save MTurk df of approved responses and df of duplicate responses.

```{r save-dfs-bm}
write.csv(dfm_b_app, "./prepped-data/sep-dfs/b-mturk-app-data0815.csv")
write.csv(dfm_b_dup, "./prepped-data/sep-dfs/b-mturk-dup-data0815.csv")
```

## Qualtrics

Excluding:

* Unapproved
* Duplicates
* Unfinished
* Did not consent
* Under 18

```{r ex-b}
dfq_b1 <- dfq_b %>% 
  filter(id %in% dfm_b_app$id) # filter out unapproved

dfq_b1 <- dfq_b1 %>% 
  filter(!id %in% dfm_b_dup$id) # filter out duplicates

dfq_b1 <- dfq_b1 %>% 
  filter(finished == 1) # filter out unfinished

dfq_b1 <- dfq_b1 %>% 
  filter(consent == 1) # filter out no consent

dfq_b1 <- dfq_b1 %>% 
  filter(dem_age >= 18 & dem_age != 4) # only Ss 18 or older (for some reason it
# won't exclude this person who reported their age as 4 unless I specify it)
```

Renaming variable names to distinguish baseline data from study data.

```{r col-names-b}
bcols <- dfq_b1 %>%
  colnames()

bcols1 <- rep(NA, length(bcols))

for (i in 1:length(bcols)) {
  
  bcols1[i] = paste(bcols[i], "b", sep = "_")

}

colnames(dfq_b1) <- bcols1 # change col names to indicate baseline

dfq_b1 <- dfq_b1 %>% 
  rename(., id = id_b) # keep id the same for joining
```

### Save dfs

Save Qualtrics df excluding duplicates and unapproved.

```{r save-df-bq}
write.csv(dfq_b1, "./prepped-data/sep-dfs/b-qual-data0815-agefixed.csv")
```

# Sci Cur Data

Import study data.

```{r import-data-sc}
# mturk
dfm_con1 <- import(here("data", "mturk-data-con1.csv")) %>% 
  clean_names() %>% 
  rename(., "id" = "amazon_identifier") %>% 
  mutate(batch = "con1") # import con HIT 1

dfm_con2 <- import(here("data", "mturk-data-con2.csv")) %>% 
  clean_names() %>% 
  rename(., "id" = "amazon_identifier") %>% 
  mutate(batch = "con2") # import con HIT 2

dfm_lib1 <- import(here("data", "mturk-data-lib1.csv"))%>% 
  clean_names() %>% 
  rename(., "id" = "amazon_identifier") %>% 
  mutate(batch = "lib1") # import lib HIT 1

dfm_lib2 <- import(here("data", "mturk-data-lib2.csv"))%>% 
  clean_names() %>% 
  rename(., "id" = "amazon_identifier") %>% 
  mutate(batch = "lib2") # import lib HIT 2

# qualtrics
dfq <- import(here("data", "qual-data.csv")) %>% 
  clean_names() # import
```

## MTurk

Bind dfs of diff HITs.

```{r bind-scm}
dfm <- rbind(dfm_con1, dfm_con2, dfm_lib1, dfm_lib2) # bind mturk data into 1 df
```

Identify, inspect, and exclude duplicate responses and unapproved responses.

```{r id-dup-scm}
dfm_dup <- dfm %>% 
  group_by(id) %>% 
  filter(n() > 1) # identify duplicates

dfm_dup %>% 
  filter(approval_status == "Approved") %>% 
  select(id)

dup_idall <- c("A3CZNWMYPM3P6E", "A2YC6PEMIRSOAA", "A2BK9RMC0NOIH8", "A2V27A9GZA1NR2",
           "ARFQ8FJRKCUZ6", "A3F9IQB0297J9D", "ATTGD1M94GFR", "A2IQ0QCTQ3KWLT",
           "A2YLK83DQ6PA5Y")

dup_id <- c("A3CZNWMYPM3P6E", "A2YC6PEMIRSOAA", "A2BK9RMC0NOIH8", "A2V27A9GZA1NR2",
           "ARFQ8FJRKCUZ6", "A3F9IQB0297J9D", "ATTGD1M94GFR", "A2YLK83DQ6PA5Y")
# A2IQ0QCTQ3KWLT does have a duplicate but they accessed the survey for 19s
# before exiting so they didn't go beyond the consent page

dfm_nodup <- dfm %>% 
  filter(!id %in% dup_id)
```

```{r id-app-scm}
dfm_app <- dfm_nodup %>% 
  filter(approval_status == "Approved") # identify approved
```

Create "key" df matching participant data w/ "batch" info.

```{r batch-sc}
dfm_batch <- dfm_app %>% 
  select(id, batch)
```

### Save dfs

Save MTurk df accounting for exclusions and batch "key" df.

```{r save-dfs-scm}
write.csv(dfm_app, "./prepped-data/sep-dfs/s1-mturk-app-data0815.csv")
write.csv(dfm_batch, "./prepped-data/sep-dfs/s1-mturk-batch-data0815.csv")
```

## Qualtrics

Identify and inspect duplicates.

```{r id-dup-scq, echo=F, eval=F}
dfq_dup <- dfq %>% 
  filter(id %in% dup_id)

dfq_dup %>% 
  group_by(id) %>% 
  summarise(n = n()) %>% 
  filter(n > 1) # identify duplicates

# taking a look to see whether can keep one of the responses
# dfq_dup %>% 
#   filter(id == "A2IQ0QCTQ3KWLT" | id == "A2YC6PEMIRSOAA" | id == "A3F9IQB0297J9D") %>% 
#   select(start_date, end_date, id, finished, duration_in_seconds, consent1,
#          bad_resp, cheat)
```

Identify responses that are outliers in duration.

```{r}
dfq <- dfq %>% 
  mutate(duration = as.numeric(duration_in_seconds),
         duration_z = scale(duration),
         duration_ex = case_when(
           duration_z >= 3 ~ "1", # z score duration times to exclude outliers
           duration_z <= -3 ~ "1",
           TRUE ~ "0"
         ))

# dfq1 %>%
#   group_by(duration_ex) %>% 
#   select(duration_in_seconds) %>% 
#   summarise(n = n())
```

Excluding:

* Duplicates
* Unfinished
* Reported cheating
* Reported bad response
* Did not consent
* Unapproved
* **Outliers**
  + **Duration (1)**

```{r ex-sc}
dfq1 <- dfq %>% 
  filter(!id %in% dup_id) # exclude duplicates

dfq1 <- dfq1 %>% 
  filter(finished == 1) # exclude unfinished

dfq1 <- dfq1 %>% 
  filter(cheat == 3) # exclude reported cheating

dfq1 <- dfq1 %>% 
  filter(bad_resp == 1) # exclude reported bad responding

dfq1 <- dfq1 %>% 
  filter(consent1 == 1) # exclude no consent

dfq1 <- dfq1 %>% 
  filter(id %in% dfm_app$id) # exclude unapproved

dfq1 <- dfq1 %>% 
  filter(duration_ex == 0) # exclude duration outliers
```

Alter column names to avoid confusion between baseline and study variables.

```{r col-names-sc}
cols <- dfq1 %>% 
  colnames()

cols1 <- rep(NA, length(cols))

for (i in 1:length(cols)) {
  
  cols1[i] = paste(cols[i], "sc", sep = "_")

} # mark study 1 cols

colnames(dfq1) <- cols1

dfq1 <- dfq1 %>% 
  rename(., id = id_sc)
```

Join study Qualtrics data with "batch" information.

```{r join-batch-sc}
dfq2 <- inner_join(dfq1, dfm_batch, by = "id", all = F)
```

Take a look at the resulting df.

```{r batch/dup-sc, echo=F, eval=F}
# dfq2 %>% 
#   filter(id %in% dfq1$id)
# 
# dfq2_unqiue <- dfq2$id %>% 
#   unique()
# 
# dfq2$id %>% 
#   unique() %>% 
#   length()
# 
# dfq1_unique <- dfq1$id %>% 
#   unique()
# 
# dfq1_unique %>% 
#   length()
# 
# dfq2 %>% 
#   filter(id %in% dfq1_unqiue)
# 
# dfq1 %>% 
#   filter(id == "A2IQ0QCTQ3KWLT")
# 
# dfm_batch$id %>% 
#   unique() %>% 
#   length()
```

### Save dfs

Save cleaned (exclusions applied) Qualtrics study df.

```{r save-dfs-scq}
write.csv(dfq2, "./prepped-data/sep-dfs/s1-qual-data0815-exdur.csv")
```

# Join (df1)

Join baseline and study Qualtrics dfs (thus excluding any Ss not in both dfs).

Excluding:

* Duplicates
* Unfinished
* Reported cheating
* Reported bad response
* Did not consent
* Unapproved
* **Was not in both dfs (b and s1)**
* Outliers
  + Duration (1)

```{r join-all}
dfj <- inner_join(dfq2, dfq_b1, by = "id", all = F) # join baseline and study 1 q data
```

## Save dfs

Save final, joined df. 

```{r save-final-df}
write.csv(dfj, "./prepped-data/joined-data0815.csv")
```
