---
title: "Does Science Curiosity Relate to Motivated Reasoning about Climate Change?"
author: Raleigh Goodwin
institute: CAIDe Seminar
date: "August 9, 2022"
output: 
  beamer_presentation:
    colortheme: "seahorse"
---

```{r include=FALSE}
knitr::opts_chunk$set(message=F, warning=F)
```

```{r lib, echo=F}
# Load libraries
library(rio)
library(here)
library(tidyverse)
library(ggplot2)
library(lme4)
library(knitr)
library(janitor)
library(Hmisc)
library(reshape2)
library(viridis)
library(kableExtra)
```

```{r data, echo=F}
# Import data
dfp_clean <- import(here("analysis-data", "cleaned-data.csv"))
dfp_prepped <- import(here("analysis-data", "prepped-for-analysis-data.csv")) %>% 
  janitor::clean_names()

dfp_lc <- dfp_prepped %>% 
  filter(conservatism_f != 4 | con != 0)
```

# Presentation Aims

\LARGE
- Taking a tour through new correlational data from a pilot study of sorts
- Questions and feedback welcome throughout!
- Also I used this as an opportunity to try my hand at making presentations in R, so bear with me in terms of aesthetics :-)

---

# Agenda

\LARGE
- Main idea and background information
- Brief overview of procedure, methods, materials
- Descriptive statistics

---

# Science Curiosity & Motivated Reasoning

\LARGE
- Science curiosity (trait-level) may mitigate selective exposure to partisan-congruent content (Kahan et al., 2017)
- Consistent with motivated reasoning theory: Curiosity is motivational and could incentivize accuracy (vs. directional) goals (Kunda, 1990)

---

# Research Question

## \LARGE Does science curiosity relate to less motivated *evaluation* of politicized science information?

## \LARGE Examining motivated reasoning through endpoint bias

Guilbeault et al. (2018) & Hardy & Jamieson (2017)

---

# Endpoint Bias & Motivated Reasoning

## Example from Guilbeault et al. (2019)

\begin{figure}
\centering
  \includegraphics[width=0.9\textwidth]{guilbeult2019.png}
\end{figure}

--- 

```{r n, echo=F}
n_total <- dfp_prepped %>% 
  nrow()

n_by_con <- dfp_prepped %>% 
  group_by(con) %>% 
  summarise(n = n(),
            percent = round(100 * (n()/nrow(dfp_prepped)), 2))

n_by_con_lc <- dfp_lc %>% 
  group_by(con) %>% 
  summarise(n = n(),
            percent = round(100 * (n()/nrow(dfp_lc)), 2))
```

```{r age, echo=F}
df_age <- dfp_prepped %>% 
  select(dem_age_b) %>% 
  psych::describe()

m_age <- df_age$mean %>% 
  round(2)
min_age <- df_age$min %>% 
  round(2)
max_age <- df_age$max %>% 
  round(2)
```

```{r gender, echo=F}
dfp_lc2 <- dfp_lc %>% 
  mutate(gender_b = as.numeric(case_when(
    dem_gender_1_b == 1 ~ 1, # man
    dem_gender_2_b == 1 ~ 2, # nb
    dem_gender_5_b == 1 ~ 3, # woman
    dem_gender_3_b == 1 ~ 4, # other
    dem_gender_4_b == 1 ~ 5 # prefer not to answer
  )))

df_gender <- dfp_lc2 %>% 
  group_by(gender_b) %>% 
  summarise(n = n(),
            percent = round(100 *(n()/nrow(dfp_lc2)),2))
```

```{r race, echo=F}
dfp_lc2 <- dfp_lc2 %>% 
  mutate(race_b = as.numeric(case_when(
    dem_race_1_b == 1 ~ 1, # white
    dem_race_2_b == 1 ~ 2, # black
    dem_race_3_b == 1 ~ 3, # hispanic
    dem_race_4_b == 1 ~ 4, # asian
    dem_race_5_b == 1 ~ 5, # nh/pi
    dem_race_6_b == 1 ~ 6, # ai
    dem_race_7_b == 1 ~ 7, # eskimo
    dem_race_8_b == 1 ~ 8, # other
    dem_race_9_b == 1 ~ 9 # prefer not to answer
  )))

df_race <- dfp_lc2 %>% 
  group_by(race_b) %>% 
  summarise(n = n(),
            percent = round(100 *(n()/nrow(dfp_lc2)),2))
```

# Participants

N$_{total}$ = `r n_total`: N$_{con}$ = `r n_by_con[3,2]`, N$_{lib}$ = `r n_by_con[1,2]`, N$_{mod}$ = `r n_by_con[2,2]`

## N$_{final}$ = `r n_total - n_by_con[2,2]`

- **Political ideology:** `r n_by_con_lc[2,3]`% conservative, `r n_by_con_lc[1,3]`% liberal
- **Age:** Mean$_{age}$ = `r m_age` (range: `r min_age`-`r max_age`)
- **Gender:** `r df_gender[3,3]`% women, `r df_gender[1,3]`% men, `r df_gender[2,3]`% non-binary folks, `r df_gender[4,3]`% preferred not to answer
- **Race:** `r df_race[1,3]`% White, `r df_race[2,3]`% Black, `r df_race[3,3]`% Hispanic, `r df_race[4,3]`% Asian

---

# Procedure

## Cohort Baseline: August 16-17, 2021

- Cognitive ability: Raven's Progressive Matrices (RPM)
- Political ideology (ANES)
- Basic demographics

## Current Study: July 13-26, 2022

- Motivated reasoning task (Guilbeault et al., 2018; Hardy & Jamieson, 2017)
- Science curiosity scale (Kahan et al., 2017)

---

# Measures: Cognitive Ability

## Raven's Progressive Matrices

\begin{figure}
  \centering
    \includegraphics[width=0.7\textwidth]{ravens.png}
 \end{figure}
 
---

# Measures: Science Curiosity

## Adapted from Science Curiosity Scale (Kahan et al., 2017)

\begin{figure}
  \centering
    \includegraphics[width=0.7\textwidth]{scs-example.png}
 \end{figure}

---

# Measures: Motivated Reasoning

## Think about 20 years from now...

\begin{figure}
  \centering
    \includegraphics[width=.8\textwidth]{endpoint-plot.png}
 \end{figure}
 
---

# Measures: Motivated Reasoning

## Coding Endpoint Bias

\begin{figure}
  \centering
    \includegraphics[width=1\textwidth]{endpoint-prompt.png}
 \end{figure}

---

```{r echo=F}
scrpm_tab <- dfp_lc2 %>% 
  mutate(RPM = rav_scored,
         SciCuriosity = sc_scored,
         ArcticIce = arctic20,
         GlobalTemp = temp20,
         OzoneHole = ozone20,
         AirQuality = airqual20,
         CO2 = co220,
         IceSheets = icesheets20,
         BacteriaPop = diatho20,
         EarthquakeMag = quake20) %>% 
  select(SciCuriosity, RPM, ArcticIce, GlobalTemp, OzoneHole, AirQuality, IceSheets, BacteriaPop, EarthquakeMag) %>% 
  psych::describe() %>% 
  mutate(Min = min,
         Max = max,
         Mean = round(mean, 2),
         SD = round(sd, 2),
         Median = median,
         Skew = round(skew, 2)) %>% 
  as.data.frame() %>% 
  select(Min, Max, SD, Mean, Median, Skew) %>% 
  kable()
```

# Basic Descriptives

`r scrpm_tab`

---

# Basic Descriptives

## Correlations

```{r echo=F, eval=F}
rcorr(dfp_lc2$sc_scored, dfp_lc2$rav_scored)

sc_1 <- lm(sc_scored ~ 1, dfp_lc2)
summary(sc_1)
scrpm <- lm(sc_scored ~ 1 + rav_scored, dfp_lc2)
summary(scrpm)
anova(sc_1, scrpm)
```

```{r echo=F}
dfp_corr <- dfp_lc2 %>% 
  select(conservatism_f, sc_scored, rav_scored, arctic20, temp20, ozone20, airqual20, co220, icesheets20, diatho20, quake20) %>% 
  mutate(Con = conservatism_f,
         SciCur = sc_scored,
         RPM = rav_scored,
         Arctic = arctic20,
         Temp = temp20,
         Ozone = ozone20,
         AirQual = airqual20,
         CO2 = co220,
         IceSheets = icesheets20,
         Bacteria = diatho20,
         Quake = quake20) %>% 
  select(-contains("20"), -contains("scored"), -contains("_f"))

get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }

cormap <- cor(dfp_corr) %>% 
  get_upper_tri() %>% 
  melt() %>% 
  ggplot(aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_viridis(limits = c(-1,1), discrete=FALSE, option = "A",
                     name="Pearson\nCorrelation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25, vjust = 1, 
    size = 13, hjust = 1)) +
  coord_fixed() +
  geom_text(aes(Var1, Var2, label = round(value, 2)), color = "white", size = 3) +
  theme(
    axis.text.y = element_text(size = 13),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
  panel.background = element_blank(),
  panel.grid.major = element_blank())
```

\begin{figure}
  \centering
    \includegraphics[width=1\textwidth]{cormap.png}
 \end{figure}

\small Conservatism ("Con"): "Extremely liberal" (1) to "Extremely conservative" (7)

---

# Descriptives by Ideology

```{r echo=F}
desc_mr_var_names <- dfp_lc2 %>% 
  mutate(RPM = rav_scored,
         SciCuriosity = sc_scored,
         ArcticIce = arctic20,
         GlobalTemp = temp20,
         OzoneHole = ozone20,
         AirQuality = airqual20,
         CO2 = co220,
         IceSheets = icesheets20,
         Bacteria = diatho20,
         Earthquakes = quake20,
         Con = con)

desc_mr_con <- desc_mr_var_names %>% 
  filter(Con == 1) %>% 
  select(SciCuriosity, RPM, ArcticIce, GlobalTemp, OzoneHole, AirQuality, IceSheets, Bacteria, Earthquakes) %>%
  psych::describe()%>% 
  mutate(C_Min = min,
         C_Max = max,
         M_c = round(mean, 2),
         SD_c = round(sd, 2),
         C_Median = median,
         Skew_c = round(skew, 2)) %>% 
  as.data.frame() %>% 
  select(M_c, SD_c, Skew_c)

desc_mr_lib <- desc_mr_var_names %>% 
  filter(Con == -1) %>% 
  select(SciCuriosity, RPM, ArcticIce, GlobalTemp, OzoneHole, AirQuality, IceSheets, Bacteria, Earthquakes) %>%
  psych::describe()%>% 
  mutate(L_Min = min,
         L_Max = max,
         M_l = round(mean, 2),
         SD_l = round(sd, 2),
         L_Median = median,
         Skew_l = round(skew, 2)) %>% 
  as.data.frame() %>% 
  select(M_l, SD_l, Skew_l)

desc_mr <- cbind(desc_mr_con, desc_mr_lib) %>% 
  select(contains("_c"), contains("_l")) %>% 
  kable() %>% 
  add_header_above(c(" " = 1, "Conservative" = 3, "Liberal" = 3))
```

`r desc_mr`

---

# Descriptives by Ideology

\begin{figure}
  \centering
    \includegraphics[width=1\textwidth]{contscoresplot.png}
 \end{figure}
 
---

# Endpoint Bias by Ideology

\begin{figure}
  \centering
    \includegraphics[width=1\textwidth]{mrscoresplot.png}
 \end{figure}

---

# Conclusion

- These are still very preliminary "findings"
- Still need to dig into inferential analyses (i.e., MLM)

---

# Thank you!

## \LARGE Questions, comments, or concerns?
email: raleighg@uoregon.edu
